--- 
title       : "Introduction to Instrumental Variables"
description : "This chapter will introduce you to instrumental variables (IV) analysis"


--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:9c7273586e
## The Logic of Instrumental Variables
*** =video_link
//player.vimeo.com/video/219558904

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:8a1c462927
## Visual Logic of Instrumental Variables
*** =video_link
//player.vimeo.com/video/219559011

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:8c410956e7
## Assumptions of instrumental variables
Which of the following is *not* a key identifying assumption/condition of instrumental variables?

*** =instructions
- Instrument has a causal effect on the outcome
- Instrument is randomly assigned
- Instrument has a causal effect on the treatment
- Instrument is correlated with the outcome
*** =sct
```{r}
msg1 = "Correct! The instrument cannot have a causal effect on the outcome."
msg2 = "The instrument must be randomly assigned. Try again"
msg3 = "The instrument must have a causal effect on the treatment. Try again"
msg4 = "The instrument needs to be correlated with the outcome. Try again"
test_mc(correct = 1, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:953b93e528
## Instrumental Variables in Action: Education and Wages (graphs)
*** =video_link
//player.vimeo.com/video/219559159

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:16cf760443
## Correlation and causation in instrumental variables
True or False: If you see that there is a zero correlation between the treatment variable and the outcome variable, then the treatment cannot have a causal effect on outcomes---there must be correlation between those two variables if we are to have any hope of using instrumental variables to learn about causality.

*** =instructions
- True
- False
*** =sct
```{r}
msg1 = "Try again"
msg2 = "Correct! The amount or sign of correlation between treatments and outcomes is completely irrelevant for instrumental variables analysis. That is because we are considering the case where treatments are *not* randomly assigned, so we know that correlation (or the lack thereof) does not imply causation (or the lack thereof). For instrumental variables analysis, all the assumptions are about how the instrument itself relates to the other variables."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2))
```

--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:116a9fd6fa
## Instrumental Variables in Action: Education and Wages (tables)
*** =video_link
//player.vimeo.com/video/219559305


--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:3ae8d65d96
## Some Instrumental Variables Terminology
*** =video_link
//player.vimeo.com/video/219565557


--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:bf3c9abc64
## The Problem of Weak Instruments
*** =video_link
//player.vimeo.com/video/219559996


--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:90f146a86a
## Where Do Instruments Come From?
*** =video_link
//player.vimeo.com/video/219560424


--- type:VideoExercise lang:r aspect_ratio:62.5 xp:50 skills:1 key:34850f663a
## Which Causal Effects Are We Actually Getting with Instrumental Variables?
*** =video_link
//player.vimeo.com/video/219560614


