---
title: 'The Local Average Treatment Effect of IV'
description: 'This chapter explains the treatment effect we are estimating with the instrumental variables method - LATE, or the Local Average Treatment Effect'
---

## Which Causal Effects Are We Actually Getting with Instrumental Variables?

```yaml
type: VideoExercise
key: 34850f663a
lang: r
xp: 50
skills: 1
video_link: //player.vimeo.com/video/219560614
```


---

## Defining LATE: The Local Average Treatment Effect

```yaml
type: VideoExercise
key: d8ab603518
lang: r
xp: 50
skills: 1
video_link: //player.vimeo.com/video/219560681
```


---

## LATE and the ATE for Always-Takers

```yaml
type: MultipleChoiceExercise
key: 7bd9f2f678
lang: r
xp: 50
skills: 1
```

LATE is defined as the average treatment effect for the compliers. Why don't we define and report an average treatment effect for the always takers?

`@possible_answers`
- Because we only care about compliers
- Because we already know what the treatment effect is for the always takers, by definition
- Because the always takers and the compliers have the same treatment effect
- None of the above

`@hint`


`@pre_exercise_code`
```{r}

```

`@sct`
```{r}
msg1 = "That's not necessarily true, because the always takers are also going to be treated, so we probably want to know about their outcomes as well. Try again."
msg2 = "This is not always possible, actually, since we don't have any data about when they don't take treatment (because they always get treated). Try again."
msg3 = "We would only be able to test this if we could always calculate the treatment effect for always takers, but we don't have their counterfactual outcomes (by definition), so we don't know if they have the same treatment effect as the compliers. Try again."
msg4 = "Correc, it's none of the above. The reason is that we are able to compute LATE from the data, (i.e. we can show it is 'identified'.) But the average treatment effect for always takers is not nearly as easy to get, or indeed could be impossible to learn: a necessary condition to calculate treatment effects is to observe outcomes for individuals in a population who do and don’t receive treatment. Since always takers always get treated, we don’t have any data about when they don’t take treatment. So we focus on LATE not so much because of intrinsic interest, although that is often the case, but also because it is something we are able to learn from the data."
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---

## The "No Defiers" Assumption

```yaml
type: VideoExercise
key: 44c9ebbb15
lang: r
xp: 50
skills: 1
video_link: //player.vimeo.com/video/219560834
```


---

## Computing LATE, Part 1: Dividing the Population

```yaml
type: VideoExercise
key: 1606faf89a
lang: r
xp: 50
skills: 1
video_link: //player.vimeo.com/video/219561008
```


---

## Computing LATE, Part 2: Finding the Compliers

```yaml
type: VideoExercise
key: d64d8a91b4
lang: r
xp: 50
skills: 1
video_link: //player.vimeo.com/video/219561095
```


---

## Computing LATE, Part 3: Getting a Result

```yaml
type: VideoExercise
key: a25130be8d
lang: r
xp: 50
skills: 1
video_link: //player.vimeo.com/video/219561285
```


---

## The Pros & Cons of LATE

```yaml
type: VideoExercise
key: edfd010726
lang: r
xp: 50
skills: 1
video_link: //player.vimeo.com/video/219561575
```


---

## Interpreting Different Results when Working with Noncompliance

```yaml
type: MultipleChoiceExercise
key: ccb5f3da17
lang: r
xp: 50
skills: 1
```

Suppose you run an experiment where individuals are randomly selected to receive a voucher for free healthcare (as in the Oregon experiment), but only half of the recipients use the voucher. You suspect that the choice to take use the voucher may be related to your health status. To deal with noncompliance but still find a causal effect, you try all four approaches: bounds analysis, assuming random compliance, intention-to-treat analysis, and instrumental variables. That gives you four results. What should you expect to see?

`@possible_answers`
- All the results will be the identical, regardless of the method
- Some of the results will be the identical, but some will be different
- None of the results will be the identical
- It doesn’t matter because only the result from instrumental variables analysis is credible

`@hint`


`@pre_exercise_code`
```{r}

```

`@sct`
```{r}
msg1 = "Remember that bounds analysis only gives you a range of values, but the other three methods will provide a single point estimate, so they can't all be identical. Try again."
msg2 = "Even if we get answers that are statistically similar (within each other's confidence intervals), the interpretations of each result are slightly different depending on the method. Try again."
msg3 = "Correct! First, the bounds analysis did not require any assumptions, so the true ATE will be within the bounds we derived there. And remember that bounds analysis always gives, well, a bound! And as long as there is some noncompliance, that bound will always be positive. So it can’t possible give the same result as using IV, assuming random compliance, or doing intention-to-treat analysis. That’s because all three of those methods deliver a single point. Moreover, each of those methods will typically deliver different answers. Which one is more credible depends on what you think about the assumptions used to compute those numbers."
msg4 = "Instrumental variables are not the best method just because they are hard! Try again."
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---

## ATEs, CATEs, and LATEs: What's the Difference?

```yaml
type: VideoExercise
key: 162436a553
lang: r
xp: 50
skills: 1
video_link: //player.vimeo.com/video/219561649
```


---

## An Alternate Way to LATE?

```yaml
type: MultipleChoiceExercise
key: 0311238e09
lang: r
xp: 50
skills: 1
```

In this video, we saw many different ways to divide up the population of units. So, when I get my dataset, all I need to do is divide my units up into groups like in this module, and then I can compute the corresponding treatment effects by just taking the average outcome within each group.

`@possible_answers`
- True
- False
- Maybe

`@hint`


`@pre_exercise_code`
```{r}

```

`@sct`
```{r}
msg1 = "Remember that in reality, we don't get data with units divided so clearly into these neat groups. We can't actually tell which units are in each group, so we can't just calculate averages. Try again."
msg2 = "Correct, it's definitely false! First, we know from very early on in these videos that learning causality is rarely as easy as just computing a single average. At the least we need to compute the *difference* in two averages, like in a randomized experiment. Beyond that, however, we have discussed in other videos that we do not actually know who is or is not a complier in our dataset, for example. So, in our data, we cannot actually split people up as seen in this module. The split we discussed here is done as if we knew things about people which we cannot possibly know with an actual dataset."
msg3 = "There is a clear answer to this question. Try again."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3))
```
